---
title: "Lab 2"
author: 
  - Makayla Whitney
  - Joe Swinehart
  - Janette Avelar
  - David Fainstein
date: "1/27/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r raw data}
if(!require(fivethirtyeight)) install.packages("fivethirtyeight", repos = "http://cran.us.r-project.org")
remotes::install_github("fivethirtyeightdata/fivethirtyeightdata")
library(fivethirtyeight)
library(fivethirtyeightdata)
library(here)
library(tidyverse)
library(lubridate)
library(waffle)
```

Visualize the change in trends for each hurricane over time in one plot with three scales. Map color to a discrete scale.
```{r google trends viz 1}

trends_clean <- google_trends %>% 
  pivot_longer(c(`hurricane_harvey_us`, `hurricane_irma_us`, `hurricane_maria_us`, `hurricane_jose_us`), names_to = "Hurricane", values_to = "Ranking")

ggplot(data=trends_clean, aes(x=date, y=Ranking, color = Hurricane))+
geom_line()+ 
  scale_color_discrete(name = "Hurricane", labels = c("Harvey", "Irma", "Jose", "Maria"))+
  labs(title = "Google Search Popularity", 
       subtitle = "According to the Google Trends dashboard",
       x = "Date",
       y = "Search Ranking")
  

```

Visualize the trends again, this time with solid shapes.
```{r google trends viz 2}
#Geom_ribbons have updated ymin & ymax so they appear on the figures
#thanks, David...I couldn't figure it out!
trends_clean %>% 
  ggplot(aes(date, Ranking)) + 
  geom_line()+
  geom_ribbon(aes(
    ymin = 0, 
    ymax = Ranking), 
    fill = "cornflowerblue",
    alpha = .85)+
  facet_wrap(~Hurricane, ncol = 1) +
  theme_minimal()+
  labs(title = "Google Search Popularity", 
       subtitle = "According to the Google Trends dashboard",
       x = "Date",
       y = "Search Ranking")
  

```

Same relationship, but mapped to a continuous scale.

```{r google trends viz 3}

  ggplot(trends_clean, aes(date, Hurricane)) + 
  geom_tile(aes(fill=Ranking)) +
  scale_fill_viridis_c(option = "plasma") +
  theme_minimal()+
  labs(title = "Google Search Popularity", 
       subtitle = "According to the Google Trends dashboard",
       x = "Date",
       y = "Hurricane")

```

```{r google trends viz 4, warning=FALSE}
#Creating date and hurricane vectors to form landfall data frame
  date <- as.Date(c('2017-08-25', '2017-09-10', NA, '2017-09-20'))
  Hurricane <- c('hurricane_harvey_us', 'hurricane_irma_us', 'hurricane_jose_us', 'hurricane_maria_us')
#Binding together vectors into data frame
  hurricane_landfall <- data.frame(date, Hurricane)
#Viz 1 with vertical lines and labels marking landfall
  ggplot(trends_clean, aes(date, Ranking, color = Hurricane)) +
    geom_line(size = 1) +
    scale_color_discrete(name = "Hurricane", labels = c("Harvey", "Irma", "Jose", "Maria"))+
    geom_vline(aes(xintercept = date, color = Hurricane), 
               hurricane_landfall, 
               size = 1,
               alpha = .5,
               linetype = "longdash") +
    geom_text(aes(x = as.Date("2017-08-25"), y = 50, label = "Landfall\nAug 25"), size = 4, color = "darkgray") +
    geom_text(aes(x = as.Date("2017-09-10"), y = 90, label = "Landfall\nSep 10"), size = 4, color = "darkgray") +
    geom_text(aes(x = as.Date("2017-09-18"), y = 30, label = "Landfall\nSep 18"), size = 4, color = "darkgray") +
    #geom_label(aes(label = Hurricane)) + <----- #This freaking line won't work ignore it
    theme_minimal() +
    labs(title = "Google Searches of Hurricanes - 2017", 
        caption = "Source: Google Trends dashboard",
        x = "Date",
        y = "Search Ranking")

```

Replicate the “National cable news networks” plot from the story using the tv_states data. Don’t worry about all the labels. Try to match the colors and worry/obsess about getting them exact.

```{r tv states viz}
#Near replication of the 538 figure
ggplot(tv_states) +
#Placing lines first
  geom_line(aes(x = date, y = florida), color = "#FFFFFF") +   
  geom_line(aes(x = date, y = texas), color = "#FFFFFF") + 
  geom_line(aes(x = date, y = puerto_rico), color = "#FFFFFF") +  
#Filling in ribbons second  
  geom_ribbon(aes(x = date, y = florida, ymin = 0, ymax = florida), fill = "#F55285", alpha = 0.98) +   
  geom_ribbon(aes(x = date, y = texas, ymin = 0, ymax = texas), fill = "#ED723C", alpha = 0.98) + 
  geom_ribbon(aes(x = date, y = puerto_rico, ymin = 0, ymax = puerto_rico), fill = "#40C1C9", alpha = 0.98) +  
#Identifying incidents (e.g., hurricane landfalls)
  geom_text(aes(x = as.Date("2017-08-25"), y = 3.1, label = "Harvey\nLandfall"), size = 4, color = "grey") +
  geom_vline(xintercept = as.Date("2017-08-25"), linetype = "dotted", color= "grey", size = 0.75) +
  geom_text(aes(x = as.Date("2017-09-07"), y = 3.3, label = "Irma\nLandfall"), size = 4, color = "grey") +
  geom_vline(xintercept = as.Date("2017-09-07"), linetype = "dotted", color= "grey", size = 0.75) +
  geom_text(aes(x = as.Date("2017-09-18"), y = 2.1, label = "Maria\nLandfall"), size = 4, color = "grey") +
  geom_vline(xintercept = as.Date("2017-09-18"), linetype = "dotted", color= "grey", size = 0.75) +
  geom_text(aes(x = as.Date("2017-10-01"), y = 2.7, label = "Las Vegas\nshooting"), size = 4, color = "grey") +
  geom_vline(xintercept = as.Date("2017-10-01"), linetype = "dotted", color= "grey", size = 0.75) +
#Labeling each state  
  geom_text(aes(x = as.Date("2017-08-28"), y = 0.8, label = "Texas"), size = 5, color = "black") +
  geom_text(aes(x = as.Date("2017-09-09"), y = 1.5, label = "Florida"), size = 5, color = "black") +
  geom_text(aes(x = as.Date("2017-10-01"), y = 1.35, label = "Puerto Rico"), size = 5, color = "black") +  
  labs(x = "",
       y = "Share of sentences",
       title = "National Cable News Networks") +
  theme_minimal()

```

Use the comic_characters dataset to visualize the proportion of different genders using at least three different methods. With your final visualization, make any additional modifications you think would help reduce the cognitive load necessary to comprehend the figure.

```{r comic characters viz 1}

genderplot1 <- comic_characters %>%
  select("publisher",
         "name",
         "sex",
         "alive",
         "appearances")
#%>%
  #mutate(appearances = as.character(appearances),
        #(fct_reorder(sex, "Male Characters",
                      #"Female Characters",
                     #"Genderless Characters",
                     #"Agender Characters",
                    # "NA" = "No Gender")))

genderplot1final <- genderplot1 %>%
  mutate(sex = factor(sex,
                              levels = c("Male Characters",
                                         "Female Characters",
                                         "NA",
                                         "Genderless Characters",
                                         "Agender Characters",
                                         "Genderfluid Characters",
                                         "Transgender Characters")))

ggplot(genderplot1final, aes(y = sex)) +
  geom_bar(aes(y = sex, fill = sex)) +
  scale_colour_viridis_d() +
  labs(title = "Proportion of Characters by Sex", 
       subtitle = "As appear within DC or Marvel Comics",
       x = "Number of Appearances",
       y = "Gender Identification")

```

```{r comic characters viz 2 - FINAL VIZ, warning = FALSE}
#Prep data for plot
genderplot2 <- comic_characters %>%
  select("publisher",
         "name",
         "sex",
         "alive",
         "appearances") %>% 
  drop_na() %>% 
  mutate(sex = fct_recode(sex,
                          "Male" = "Male Characters", 
                          "Female" = "Female Characters",
                          "Agender" = "Agender Characters",
                          "Genderless" = "Genderless Characters",
                          "Genderfluid" = "Genderfluid Characters",
                          "Transgender" = "Transgender Characters"))

#Leveraged Makayla's existing code to summarise then sort  
ordered_gender <- genderplot2 %>% 
  group_by(sex) %>% 
  summarise(appearances = sum(appearances)) %>% 
  arrange(desc(appearances))

#Still can't quite connect the calculated and sorted sum list with the whole dataset, so I wrote
#it out by hand *sigh*
genderplot2$sex <- factor(genderplot2$sex, levels = c("Male", "Female", "Agender", "Genderless", "Genderfluid", "Transgender"))

#Plot
  ggplot(genderplot2, aes(x = sex, y = appearances, fill = sex)) +
    geom_jitter(aes(color = sex, width = .25), show.legend = FALSE, size = .65, stat = "identity") +
    scale_colour_viridis_d() +
    labs(title = "DC and Marvel Comic Characters by Gender Identification",
       x = "",
       y = "") +
    theme_dark() +
    theme(
      plot.title = element_text(family = "Times", face = "bold", size = 18, lineheight = .8),
      axis.text.x = element_text(face = "bold", hjust = 1),
                    element_line(linetype = 0),
      axis.text.y = element_blank())
    #annotate(geom_label())


#To Do:  
# - X-axis spacing
# - Add labels with n?
  
```

```{r comic characters viz 3}
#If you don't have this package, then here's the install line:
#devtools::install_git("https://git.rud.is/hrbrmstr/waffle.git")

#Prepping the waffle
genderplot3 <- comic_characters %>% 
  select(name, sex, appearances) %>% 
  group_by(sex) %>% 
  filter(!is.na(sex)) %>% 
  summarize(n = n()) %>% 
  mutate(percent = (n/sum(n))*100)
  
#Specifying variable for waffle
sex_counts <- genderplot3$percent
names(sex_counts) <- genderplot3$sex
#The waffle
waffle(sex_counts)

```

Using a waffle chart theoretically sounded great, but as in Viz 1, the data did not present properly.

Mainly, the numbers for non-male or -female characters were so low that they didn't show up in the viz.

